An Efficient Minibatch Acceptance Test for Metropolis-Hastings

### 2016-09-23 nips_2016.tex
Updated Section 6.1

We have the additional constraint that $u_j > 0$ for all $j$, since $u$
represents a density. We first explored optimizing this problem with linear
programming to find a $u$ minimizing $|Mu - v|$ subject to $u \ge 0$. This was
tractable up to a few hundred dimensions for $u$. However, the discretization
error is bounded by the curvature of the underlying functions which are here
slightly less than one, i.e.  the errors are of the order of $(V/N)^2$.
{\color{blue} Daniel: unfortunately, I do not understand what this means (i.e.,
the curvature and how the $(V/N)^2$ is obtained).} Here we chose $V=20$ to
provide adequate containment of the distributions (the CDFs are extremely close
to either zero or one outside this range). So with 200 points, we have a
discretization error of approximately 0.01, which is relatively large.  To yield
higher resolution and lower error, we switched to a least squares solution:
\begin{equation}\label{eq:optimization_l2}
    u^* = \arg_u \min \; \|Mu-v\|_2^2 + \lambda \|u\|_2^2,
\end{equation}
where we add the standard regularization factor $\lambda$ for stability in high
dimensions. The solution is well-known from the normal equations: $u^* = (M^TM +
\lambda I)^{-1}M^Tv$, and in practice, it yields an acceptable $L_1$ norm for
Equation~\ref{eq:optimization_l1}.

@@

To start, we use the following quantitative central-limit result (from
\cite{explicit-clt05}):

@@

PS: I'd recommend not putting sample mean and sample variance in an equation if
the $\sigma(\hat{X})$ gets clarified, since those computations are common
knowledge.}

@@

To apply this to our test, let $X_i = \Lambda_i(\theta,\theta') -
\Lambda(\theta,\theta')$, so that the $X_i$ are zero-mean, i.i.d. variables. If
instead of all $n$ samples, we only extract a subset of $b$ samples
corresponding to our minibatch, we can connect $\hat{X}$ with our $\Delta^*$
term:
\begin{equation}\label{eq:x-hat}
    \hat{X} = \Delta^*(\theta,\theta') - \Delta(\theta,\theta'),
\end{equation}
so that $\sigma(\hat{X}) = \sigma(\Delta^*)$. This results in the following
Corollary:

@@

\begin{proof}
We apply Lemma~\ref{lem:cdf_bounds} twice. First take:
\begin{equation}
    P(y) = {\rm Pr}(\Delta^* < y) \qquad {\rm and}\qquad Q(y) = \Phi\left(\frac{y-\Delta}{\sigma(\Delta^*)}\right)
\end{equation}
and convolve with the distribution of $X_n$ which has density $\phi(X/\sigma_n)$
where $\sigma_n^2 = 1 - \sigma^2(\Delta^*)$. This yields the next iteration of
$P$ and $Q$
\begin{equation}
    P'(y) = {\rm Pr}(\Delta^*+X_{\rm norm} < y) \qquad {\rm and}\qquad Q'(y) = \Phi\left({y-\Delta}\right)
\end{equation}
Now we convolve with the distribution of $X_{\rm corr}$ which gives:
\begin{equation}
    P''(y) = {\rm Pr}(\Delta^*+X_{\rm norm}+X_{\rm corr} < y) \qquad {\rm and}\qquad Q''(y) = S\left({y-\Delta}\right)
\end{equation}
Both steps preserve the error bound $\epsilon(\theta,\theta',b)$. Finally
$S(y-\Delta)$ is a logistic CDF centered at $\Delta$, and so $S(y-\Delta) = {\rm
Pr}(\Delta + X_{\rm log} < y)$ for a logistic random $X_{\rm log}$. We conclude
that the probability of acceptance for the actual test ${\rm Pr}(\Delta^*+X_{\rm
norm}+X_{\rm corr} > 0)$ differs from the exact test ${\rm Pr}(\Delta+X_{\rm
log} > 0)$ by at most $\epsilon$.
\end{proof}

@@

In fact we can do better than this approximation (showing the error decreases as
$O(n^{-1})$) by using a more precise limit distribution under an additional
assumption. Let $\mu_i$ denote the $i^{th}$ moment, and $b_i$ denote the
$i^{th}$ absolute moment of $X$. If Cramer's condition holds:
\begin{equation}\label{eq:cramers_condition}
    \lim_{t \to \infty} \sup |E(\exp(i t X))| < 1,
\end{equation}
then Equation 2.2 in Bentkus et al.'s work on Edgeworth
expansions~\cite{Bentkus97} provides the following Lemma:

@@

Here the limit $G_n(x,y)$ is a normal CDF plus a correction term that decays as
$n^{1/2}$. Importantly, since $x^2\phi(x) = \phi^{''}(x) + \phi(x)$ where
$\phi(x)=\Phi'(x)$, the correction term can be rewritten giving:
\begin{equation}\label{eq:GNderivatives}
    G_n(x,y) = \Phi(x) + \frac{y}{6\sqrt{n}}(2\phi^{''}(x)+3\phi(x))
\end{equation}
From which we see that $G_n(x,y)$ is a linear combination of $\Phi(x)$,
$\phi(x)$ and $\phi^{''}(x)$. This is quite fortuitous. In Algorithm 1, we
correct for the difference in $\sigma$ between $\Delta^*$ and the variance
needed by $X_{\rm corr}$ using $X_{\rm norm}$. This same method works when we
wish to estimate the error in $\Delta^*$ vs $G_n(x,y)$. Since all of the
component functions of $G_n(x,y)$ are derivatives of a (unit variance)
$\Phi(x)$, adding a normal variable with variance $\sigma'$ increases the
variance of all three functions to $1+\sigma'$. Thus we add $X_{\rm norm}$ as
per Algorithm 1 preserving the limit in Equation~\ref{eq:GNderivatives}.

The deconvolution approach can be used to construct a correction variable
$X_{\rm corr}$ between $G_n(x,y)$ and $S(x)$ the standard logistic function. An
additional complexity is that $G_n(x,y)$ has additional parameters $y$ and $n$.
Since these act as a single multiplier $\frac{y}{6\sqrt{n}}$ in
Equation~\ref{eq:GNderivatives}, its enough to consider a function $g(x,y')$
parametrized by $y'= \frac{y}{6\sqrt{n}}$. This function can be computed and
saved offline. As we have shown above, errors in the ``limit'' function
propagate directly through as errors in the acceptance test.  To achieve a test
error of say $1e-6$ (close to single floating point precision), we need a $y'$
spacing of $1e-6$. It should not be necessary to tabulate values all the way to
$y'=1$, since $y'$ is scaled inversely by the square root of minibatch size.
Assuming a max $y'$ of 0.1 requires us to tabulate about 100,000.  Since our $x$
resolution is 10,000, this leads to a table with about 1 billion values, which
can comfortably be stored in memory.  However, if $g(x,y)$ is moderately smooth
in $y$, it should be possible to achieve similar accuracy with a much smaller
table. We leave further analysis and experiments with $g(x,y)$ as future work.

%% \textbf{The Convolution}. Figure~\ref{fig:deconvolution} demonstrates
%% Equation~\ref{eq:deconvolution} and provides three (color-coded) examples of densities for $X_{\rm
%% norm}$ and $X_{\rm corr}$. The density of $X$, a logistic random variable with mean $\mu = 0$ and
%% scale $s=1$, is known, as are the $X_{\rm norm}$ densities for different standard deviations. The
%% deconvolution provides us with the $X_{\rm corr}$ distributions. Note, however, that as the standard
%% deviation of $X_{\rm norm}$ increases, the $X_{\rm corr}$ distribution becomes increasingly unstable
%% and ``bumpier'', because the logistic curve has fatter tails than Gaussians. In addition, we cap
%% the standard deviation possibilities of $X_{\rm norm}$ at $\approx 1.2$ in our experiments. In our
%% experiments, we discretize the possible standard deviations of $X_{\rm norm}$, and we also limit the
%% densities to consider the range $[-10,10]$ as shown in Figure~\ref{fig:deconvolution}.

%% \textbf{Variance Preconditioning}. Returning to the discussion from Section~\ref{sec:introduction},
%% our test requires a variance check. Equation~\ref{eq:deconvolution} means that the variance of $X$,
%% which is $\pi^2/3\approx 3.29$ is an upper bound on the variances of $X_{\rm norm}$. Therefore, the
%% test cannot work if the variance of $X_{\rm norm}$ is too large. The standard deviation of $X$
%% is about $\sqrt{3.29}\approx 1.81$, so this bounds ${\rm std}(X_{\rm norm})$ (as well as ${\rm
%% std}(\Delta')$). From the discussion above, we bound the standard deviation at 1.2 so that $X_{\rm
%% corr}$ can handle the remaining noise. Recall from Section~\ref{ssec:deltas_minibatch} that we
%% estimate the ${\rm std}(X_{\rm norm})$ each iteration. If the variance/standard deviation test
%% fails, then we can (a) skip the iteration, (b) increase the minibatch size, (c) change the proposal
%% to decrease step sizes, or (d) increase temperature, as we now discuss.

%% \textbf{Temperature}. One way to satisfy the variance precondition is to increase the temperature of
%% the target distribution. For a temperature $T>1$, the augmented target distribution would become
%% \begin{equation}\label{eq:log_temperature}
%% \log p(\theta \mid x_1,\ldots,x_N) \approx \log p(\theta) + \frac{1}{T}\frac{N}{n} \sum_{i=1}^n\log p(x_i \mid \theta).
%% \end{equation}
%% with an extra $1/T$ term augmented. As the temperature increases, the values of $\Delta'$ get
%% smaller in absolute value, thus reducing variance. The flatter posterior results in easier mixing.

@@

We start with a simple Gaussian mixture model, borrowing an experiment from~\cite{langevin_2011}.
The parameter is 2-D, $\theta = (\theta_1,\theta_2)$, and the parameter/data generation process is
\begin{equation}\label{eq:data_generation}
(\theta_1,\theta_2) \sim \mathcal{N}((0,0), {\rm diag}(\sigma_1^2, \sigma_2^2));\quad \quad x_i \sim
0.5 \cdot \mathcal{N}(\theta_1, \sigma_x^2) + 0.5 \cdot \mathcal{N}(\theta_1+\theta_2, \sigma_x^2).
\end{equation}
We set $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$. Fixing $\theta = (0,1)$, we draw 1,000,000
data points so that the target distribution is $p(\theta)\prod_{i=1}^{1,000,000}p(x_i\mid \theta)$, with
the prior based on the $\theta$ generation process in Equation~\ref{eq:data_generation}. This
results in rather sharp posterior modes and high ${\rm std}(\Delta')$ estimates, so we apply a
temperature $T=10,000$ to reduce ${\rm std}(\Delta')$. Taking logs, we get the target as shown in the
far left of Figure~\ref{fig:gauss_mix_1}.

We run MCMC with our MH test using minibatch size $m=100$. %We also run this using standard
%full-batch MCMC (``standard MCMC'') with the standard test from Equation~\ref{eq:traditional}, 
We also run this using the method from~\cite{cutting_mh_2014} (``Cut MH'')  and the method from 
~\cite{icml2014c1_bardenet14}. For the method in \cite{cutting_mh_2014}, we use $m=100$ and
increment minibatches by that amount within a test if necessary. As for the method in ~\cite{icml2014c1_bardenet14}, we use $m=100$ and increase the minibatch size geometrically with a ratio of $\gamma = 1.5$. The tolerance for making a
decision in \cite{cutting_mh_2014} is $\epsilon=0.005$, and the error bound control parameter in ~\cite{icml2014c1_bardenet14} is $p = 2, \delta = 0.01$. To make comparisons easier, all three use the same random walk proposer with covariance $\Sigma={\rm diag}(0.3, 0.3)$. This is a poor proposer, but it is sufficient for
our purposes as the quality of the samples will be primarily due to the MH test. All methods are run
5000 times to collect 5000 samples.

\begin{figure}[t]
    \centering
    % \includegraphics[width=1\linewidth]{cloud_v01.png}
    \includegraphics[width=1\linewidth]{GaussianMixtureResult/posterior_of_gaussian.png}
    \caption{
    The log posterior contours and scatter plots of sampled $\theta$ values using different methods. 
    }
    \label{fig:gauss_mix_1}
    %\vspace{-10pt}
\end{figure}

\begin{figure}[t]
    \centering
   % \includegraphics[width=1\linewidth]{adaptive_and_ours_information_v01.png}
   \includegraphics[width=1\linewidth]{GaussianMixtureResult/minibatch_size_gaussian.png}
    \caption{
    Histograms of minibatch sizes %and ${\rm std}(\Delta')$ values, along with ``${\rm std}(\Delta')$
   % violations'' (right).
    }
    \label{fig:gauss_mix_2}
    %\vspace{-10pt}
\end{figure}

\begin{figure}[t]
	\centering
	\includegraphics[width=\linewidth]{GaussianMixtureResult/first_second_part_gaussian.png}
	\caption{Value of first and second part of equation (9) in ~\cite{icml2014c1_bardenet14} with respect to actual minibatch size}
	\label{fig:gauss_bardenet}
\end{figure}

%\begin{figure}[t]
    %\centering
    %\includegraphics[width=1\linewidth]{CI_v01.png}
   % \caption{
   % Histograms of ${\rm std}(\Delta')$, confidence interval of ${\rm std}(\Delta')$ (middle), along with $\rho / \sigma^3$ (right).
    %}
   % \label{fig:gauss_mix_3}
    %\vspace{-10pt}
%\end{figure}

Figure~\ref{fig:gauss_mix_1} shows scatter plots of the resulting $\theta$ samples for the four
methods, with darker regions indicating a greater density of points. The three methods obtain the
same rough form of the posterior, so our MH test can indeed result in the same posterior as the
other two methods. %However, the fourth method (i.e. just treat the $\Delta'$ which comes from minibatch as $\Delta$ which comes from the complete data) reveals wider spread distribution than the posterior. From Figure~\ref{fig:gauss_mix_1} we illustrate our proposed method can obtain the samples which have almost the same distribution as the standard MH test from the complete data, adaptive minibatch test, and the true posterior. Meanwhile, we show the benefits of our minibatch method compared with the naive minibatch method.  
Actually, based on a measure of divergence from the ground-truth distribution, our method is better than the other three methods in terms of this measurement. The measurement is to calculate the poisson likelihood of the MCMC-generated parameters over a parameter range, with $\theta_1 \in (-1.5, 2.5)$ and $\theta_2 \in (-2.5, 2.5)$. The log likelihood value for our method, cut MH method and adaptive MH method are -2144.6, -4007.4, -2598.4, respectively.

Figure~\ref{fig:gauss_mix_2} shows a histogram of the final minibatch sizes used by
the three methods in each iteration. It is obvious that our methods consumes significantly less data, and most minibatch sizes are smaller than 1000, with an average minibatch size of 420. The other two methods occasionally consumes a large proportion of the entire data set, and the average minibatch sizes are 16378, 16755 for cut MH method, adaptive MH method, respectively. In terms of average minibatch size, our method is nearly 39 times faster the cut MH method, and nearly 40 times faster than the adaptive MH method.

It is useful to notice that, in~\cite{icml2014c1_bardenet14}, they need to evaluate the likelihood term $C_{\theta,\theta'}$ over the entire the data set at every iteration in order to calculate the error bound in equation (9) in ~\cite{icml2014c1_bardenet14}. In Figure ~\ref{fig:gauss_bardenet}, the value of the first part and second part of equation (9) in ~\cite{icml2014c1_bardenet14} is shown with respect to the size of actual minibatch data used. The second part which contains  the $C_{\theta,\theta'}$ term is larger than the first part, which means that the second part is non-negligible and has to be calculated at every iteration. This is one of the very time-consuming parts in this method.

%Usually, the method can make a decision with 200 samples. Occasionally, however, it must use all 10000 points, resulting in lots of wasted computation.  For this particular experiment, that happened eight times.
%The third plot in Figure~\ref{fig:gauss_mix_2} investigates the number of times ${\rm std}(\Delta')
%\ge 1.2$ for only consuming one minibatch data. For six minibatch sizes (50, 100, 150, 200, 250, and 300), we ran MCMC with
% our MH test five times and averaged the number of ``violations.'' As the size increases, the
% number of violations decreases, as expected.

\iffalse
Figure~\ref{fig:gauss_mix_3} (to the left) shows a histogram of ${\rm
  std}(\Delta')$. Figure~\ref{fig:gauss_mix_3} (to the middle) shows
the confidence interval of our estimated ${\rm std}(\Delta')$ at each
iteration \footnote{In order to show the confidence interval clearly,
  we use moving smooth to smooth the confidence interval, and plot
  each value by interval of 50 iterations}. The blue line represents
the standard deviation comes from the complete data, which can be
treated as the ``true'' ${\rm std}(\Delta')$, the green line
represents our estimation of ${\rm std}(\Delta')$ from the minibatch,
and the two red dash lines represent the confidence interval of ${\rm
  std}(\Delta')$ by adding / subtracting one ${\rm
  std(std(}\Delta'))$. This figure indicates that our estimation of
${\rm std}(\Delta')$ from one minibatch data is
accurate. Figure~\ref{fig:gauss_mix_3} (to the right) shows the value
of $\rho/\sigma^3$ at each iteration. 
\fi

>>>>

We have the additional constraint that $u_j > 0$ for all $j$, since $u$
represents a density. We first explored optimizing this problem with linear
programming to find a $u$ minimizing $|Mu - v|$ subject to $u \ge 0$. This was
tractable up to a few hundred dimensions for $u$. However, the discretization
error is bounded by the curvature of the underlying functions which are here
slightly less than one, i.e.  the errors are of the order of $(V/N)^2$.
{\color{blue} Daniel: unfortunately, I do not understand what this means (i.e.,
the curvature and how the $(V/N)^2$ is obtained).} Here we chose $V=20$ to
provide adequate containment of the distributions (the CDFs are extremely close
to either zero or one outside this range). So with 200 points, we have a
discretization error of approximately 0.01, which is relatively large.  To yield
higher resolution and lower error, we switched to a least squares solution:
\begin{equation}\label{eq:optimization_l2}
    u^* = \argmin_u\; \|Mu-v\|_2^2 + \lambda \|u\|_2^2,
\end{equation}
where we add the standard regularization factor $\lambda$ for stability in high
dimensions. The solution is well-known from the normal equations: $u^* = (M^TM +
\lambda I)^{-1}M^Tv$, and in practice, it yields an acceptable $L_1$ norm for
Equation~\ref{eq:optimization_l1}.

@@

{\color{blue} Daniel: the following lemma is, I think, from Corollary 2
in~\cite{explicit-clt05} but I cannot derive the exact lemma as we have it
here.} To start, we use the following quantitative central-limit result
(from~\cite{explicit-clt05}):

@@

PS: I'd recommend not putting sample mean and sample variance in an equation if
the $\sigma(\hat{X})$ gets clarified, since those computations are common
knowledge.

PPS: The sample mean is probably more commonly denoted as $\bar{X}$.}

@@

To apply this to our test, let $X_i = \Lambda_i(\theta,\theta') -
\Lambda(\theta,\theta')$, so that the $X_i$ are zero-mean, i.i.d. variables. If
instead of all $n$ samples, we only extract a subset of $b$ samples
corresponding to our minibatch, we can connect $\hat{X}$ with our $\Delta^*$
term:
\begin{equation}\label{eq:x-hat}
    \hat{X} = \Delta^*(\theta,\theta') - \Delta(\theta,\theta'),
\end{equation}
so that $\sigma(\hat{X}) = \sigma(\Delta^*)$. {\color{blue} Daniel: This time it
makes sense to use $\sigma(\hat{X})$ instead of $\sigma(X)$.} This results in
the following Corollary:

@@

\begin{proof}
We apply Lemma~\ref{lem:cdf_bounds} twice. First take:
\begin{equation}
    P(y) = {\rm Pr}(\Delta^* < y) \qquad {\rm and}\qquad Q(y) = \Phi\left(\frac{y-\Delta}{\sigma(\Delta^*)}\right)
\end{equation}
and convolve with the distribution of $X_n$ {\color{blue} (Daniel: to be clear I
think this was supposed to be $X_{\rm norm}$ but see comments earlier about why
I think it is not necessary.)} which has density $\phi(X/\sigma_n)$ where
$\sigma_n^2 = 1 - \sigma^2(\Delta^*)$. This yields the next iteration of $P$ and
$Q$
\begin{equation}
    P'(y) = {\rm Pr}(\Delta^*+X_{\rm norm} < y) \qquad {\rm and}\qquad Q'(y) = \Phi\left({y-\Delta}\right)
\end{equation}
Now we convolve with the distribution of $X_{\rm corr}$ which gives:
\begin{equation}
    P''(y) = {\rm Pr}(\Delta^*+X_{\rm norm}+X_{\rm corr} < y) \qquad {\rm and}\qquad Q''(y) = S\left({y-\Delta}\right)
\end{equation}
Both steps preserve the error bound $\epsilon(\theta,\theta',b)$. Finally
$S(y-\Delta)$ is a logistic CDF centered at $\Delta$, and so $S(y-\Delta) = {\rm
Pr}(\Delta + X_{\rm log} < y)$ for a logistic random $X_{\rm log}$. We conclude
that the probability of acceptance for the actual test ${\rm Pr}(\Delta^*+X_{\rm
norm}+X_{\rm corr} > 0)$ differs from the exact test ${\rm Pr}(\Delta+X_{\rm
log} > 0)$ by at most $\epsilon$.
\end{proof}

@@

In fact we can do better than this approximation (showing the error decreases as
$O(n^{-1})$) by using a more precise limit distribution under an additional
assumption. Let $\mu_i$ denote the $i^{th}$ moment, and $b_i$ denote the
$i^{th}$ absolute moment of $X$. If Cramer's condition holds:
\begin{equation}\label{eq:cramers_condition}
    \lim_{t \to \infty} \sup |E(\exp(i t X))| < 1,
\end{equation}
then Equation 2.2 in Bentkus et al.'s work on Edgeworth
expansions~\cite{Bentkus97} provides the following:

@@

Here the limit $G_n(x,y)$ is a normal CDF plus a correction term that decays as
$n^{1/2}$. {\color{blue} (Daniel: how did the equation in the next sentence
appear? I verified it and it seems correct, but it looks like it popped out of
nowhere.)} Importantly, since $x^2\phi(x) = \phi^{''}(x) + \phi(x)$ where
$\phi(x)=\Phi'(x)$, the correction term can be rewritten giving:
\begin{equation}\label{eq:GNderivatives}
    G_n(x,y) = \Phi(x) + \frac{y}{6\sqrt{n}}(2\phi^{''}(x)+3\phi(x))
\end{equation}
From which we see that $G_n(x,y)$ is a linear combination of $\Phi(x)$,
$\phi(x)$ and $\phi^{''}(x)$. This is quite fortuitous. In Algorithm 1, we
correct for the difference in $\sigma$ between $\Delta^*$ and the variance
needed by $X_{\rm corr}$ using $X_{\rm norm}$. This same method works when we
wish to estimate the error in $\Delta^*$ vs $G_n(x,y)$. Since all of the
component functions of $G_n(x,y)$ are derivatives of a (unit variance)
$\Phi(x)$, adding a normal variable with variance $\sigma'$ increases the
variance of all three functions to $1+\sigma'$. Thus we add $X_{\rm norm}$ as
per Algorithm 1 preserving the limit in Equation~\ref{eq:GNderivatives}.

The deconvolution approach can be used to construct a correction variable
$X_{\rm corr}$ between $G_n(x,y)$ and $S(x)$ the standard logistic function. An
additional complexity is that $G_n(x,y)$ has additional parameters $y$ and $n$.
Since these act as a single multiplier $\frac{y}{6\sqrt{n}}$ in
Equation~\ref{eq:GNderivatives}, its enough to consider a function $g(x,y')$
parametrized by $y'= \frac{y}{6\sqrt{n}}$. This function can be computed and
saved offline. As we have shown above, errors in the ``limit'' function
propagate directly through as errors in the acceptance test.  To achieve a test
error of say $1e-6$ (close to single floating point precision), we need a $y'$
spacing of $1e-6$. It should not be necessary to tabulate values all the way to
$y'=1$, since $y'$ is scaled inversely by the square root of minibatch size.
Assuming a max $y'$ of 0.1 requires us to tabulate about 100,000.  Since our $x$
resolution is 10,000, this leads to a table with about 1 billion values, which
can comfortably be stored in memory.  However, if $g(x,y)$ is moderately smooth
in $y$, it should be possible to achieve similar accuracy with a much smaller
table. We leave further analysis and experiments with $g(x,y)$ as future work.
{\color{blue} Daniel: I will look at the code to understand this paragraph better.}

@@

We start with a simple Gaussian mixture model, borrowing an experiment
from~\cite{langevin_2011}.  The parameter is 2-D, $\theta =
(\theta_1,\theta_2)$, and the parameter/data generation process is
\begin{equation}\label{eq:data_generation}
    (\theta_1,\theta_2) \sim \mathcal{N}((0,0), {\rm diag}(\sigma_1^2, \sigma_2^2));\quad \quad x_i \sim
    0.5 \cdot \mathcal{N}(\theta_1, \sigma_x^2) + 0.5 \cdot \mathcal{N}(\theta_1+\theta_2, \sigma_x^2).
\end{equation}
We set $\sigma_1^2 = 10, \sigma_2^2 = 1$ and $\sigma_x^2=2$. Fixing $\theta =
(0,1)$, we draw 1,000,000 data points so that the target distribution is
$p(\theta)\prod_{i=1}^{1,000,000}p(x_i\mid \theta)$, with the prior based on the
$\theta$ generation process in Equation~\ref{eq:data_generation}. This results
in rather sharp posterior modes and high ${\rm std}(\Delta')$ estimates, so we
apply a temperature $T=10,000$ to reduce ${\rm std}(\Delta')$. Taking logs, we
get the target as shown in the far left of Figure~\ref{fig:gauss_mix_1}.
{\color{blue} Daniel: the ${\rm std}(\Delta')$ stuff is outdated; let's figure
out a way to change this. We may also have to justify such a high temperature,
because this is far from the ``real'' (log) posterior.}

We run MCMC with our MH test using minibatch size $m=100$.  We also run this
using the methods from Korattikara 2014~\cite{cutting_mh_2014} and
Bardenet 2014~\cite{icml2014c1_bardenet14}. For~\cite{cutting_mh_2014}, we use
$m=100$ and increment minibatches by that amount within a test if necessary, and
for~\cite{icml2014c1_bardenet14}, we use $m=100$ and increase the minibatch size
geometrically with a ratio of $\gamma = 1.5$. The tolerance for making a
decision in~\cite{cutting_mh_2014} is $\epsilon=0.005$, and the error bound
control parameters in~\cite{icml2014c1_bardenet14} are $p = 2$ and $\delta =
0.01$. To make comparisons easier, all three use the same random walk proposer
with covariance $\Sigma={\rm diag}(0.3, 0.3)$. This is a poor proposer, but it
is sufficient for our purposes as the quality of the samples will be primarily
due to the MH test. All methods are run 5000 times to collect 5000 samples.

\begin{figure}[t]
    \centering
    \includegraphics[width=1\linewidth]{GaussianMixtureResult/posterior_of_gaussian.png}
    \caption{
    The log posterior contours and scatter plots of sampled $\theta$ values using different methods. 
    }
    \label{fig:gauss_mix_1}
\end{figure}

\begin{figure}[t]
    \centering
    \includegraphics[width=1\linewidth]{GaussianMixtureResult/minibatch_size_gaussian.png}
    \caption{
    Histograms of minibatch sizes for the three methods used in
    Section~\ref{ssec:gaussians}. {\color{blue} Daniel: The y-axis should
    probably just be ``Counts'' not ``Total Counts.'' Also, was this run with
    the \texttt{plt.tight\_layout()} command? The y-axis text for the second and
    third subplots almost overlap with the graphs to their left. Finally, I am
    still thinking about whether the y-axis should be on a log scale. I think
    the log scale is appropriate for the x-axis, it's a question of whether we
    should also apply it for the y-axis.}
    }
    \label{fig:gauss_mix_2}
\end{figure}

\begin{figure}[t]
	\centering
	\includegraphics[width=\linewidth]{GaussianMixtureResult/first_second_part_gaussian.png}
	\caption{Value of first and second part of equation (9) in~\cite{icml2014c1_bardenet14} with respect to actual minibatch size}
	\label{fig:gauss_bardenet}
\end{figure}

Figure~\ref{fig:gauss_mix_1} shows scatter plots of the resulting $\theta$
samples for the three methods, with darker regions indicating a greater density
of points. The three methods obtain the same rough form of the posterior, so our
MH test can indeed result in similar posteriors as the other two methods.

In addition, our method is superior based on the Poisson likelihood of the
MCMC-generated parameters over the range of $\theta_1 \in (-1.5, 2.5)$ and
$\theta_2 \in (-2.5, 2.5)$. This is a measure of divergence from the
ground-truth distribution. The log likelihood values for our method, Korattikara
2014, and Bardenet 2014 are -2144.6, -4007.4, and -2598.4, respectively.
{\color{blue} Daniel: I rewrote this paragraph to make it slightly less
offensive. Also, should we clarify what this means? Because to be honest, it
seems hard to tell a significant difference among the methods by looking at the
scatterplots.}

Figure~\ref{fig:gauss_mix_2} shows a histogram of the final minibatch sizes used
by the three methods in each iteration. Our methods consumes significantly less
data; most minibatch sizes are smaller than 1000, and the average size is 420.
The other two methods occasionally need to consume a large proportion of the
entire data set of one million elements; the average minibatch sizes are 16378
and 16755 for Korattikara 2014 and Bardenet 2014, respectively. {\color{blue}
Daniel: the plot says our average minibatch size is 210, but the paragraph here
says 420.}

It is useful to notice that, in~\cite{icml2014c1_bardenet14}, they need to
evaluate the likelihood term $C_{\theta,\theta'}$ over the entire the data set
at every iteration in order to calculate the error bound in equation (9)
in~\cite{icml2014c1_bardenet14}. In Figure ~\ref{fig:gauss_bardenet}, the value
of the first part and second part of equation (9)
in~\cite{icml2014c1_bardenet14} is shown with respect to the size of actual
minibatch data used. The second part which contains  the $C_{\theta,\theta'}$
term is larger than the first part, which means that the second part is
non-negligible and has to be calculated at every iteration. This is one of the
very time-consuming parts in this method.  {\color{blue} Daniel: I am not sure
if we should include this plot. Let me think about it and if we decide to
include it I will help rewrite the explanation.}
